\section{Related Work}
\label{sec:relw}

\subsection{NUMA GC Characterization}

\cite{sartor12exploring}
study JVM performance for a two-socket NUMA machine. They experiment with running
GC threads on one socket at a lower frequency than application threads on the 
other socket. In general, they find that there can be fewer GC threads than application threads, although they do not expose NUMA congestion as the 
underlying reason for this.

\subsection{Causes of Congestion}

Gidra et. al \cite{Gidra2013} analyzed the OpenJDK Parallel Scavenge GC for parallelism bottlenecks and pinpointed several contended data structures. They propose several modifications to reduce contention (and corresponding NUMA congestion).

\cite{iyengar2012scalable} study the scalability of the marking phase of the C4 algorithm \cite{Tene2011}.
They report that the duty cycles of marking phase get worse as the number of threads increases.
A primary source of this problem is the contention of work sharing in marking tasks, where multiple threads attempt to atomically update words in a side bitmap.
The JVM allocates the bitmap data structure at the initialization phase and it is likely to reside in a single NUMA node.
Therefore, the collector threads would saturate the bandwidth of that node.
In addition, the cache line is big enough to accommodate multiple words; thus, false sharing would occur with high probability.

GC work-stealing may also lead to congestion.
\cite{gidra11assessing} remarks that disabling work-stealing improves program performance for some applications.
\cite{Muddukrishna2013} suggest a locality-aware work-stealing algorithm, which calculates the distance between NUMA nodes in a system with multi-hop memory hierarchy.
An idle thread on a node attempts to steal work from the `nearest' non-empty queues.
\cite{Olivier2011} propose a hierarchical work-stealing algorithm to improve locality.
They enable one third of running threads to steal work on behalf of other threads in the same chip and push stolen work into a shared queue for local threads.
Our approach allows threads to steal from local threads only, to preserve NUMA locality.

\subsection{Reducing Congestion}

\cite{gidra15numagic} use message passing between local and remote GC 
threads to prevent NUMA congestion.

Locality improvement is the most common way to reduce congestion. This is
achieved by thread-local heaps 
\cite{marlow11multicore,domani02thread}. 
Object migration may improve locality, but causes non-trivial inter-node traffic \cite{tikir05numa}.